---
title: "Shopping Trends"
author: "Marlen Gonzalez"
date: "2023-11-27"
output: slidy_presentation
---
# Introduction
I’ve always been the consumer but never looked at shopping from the point of view from the business side of it. I don’t usually go to a store and wonder what the popular selling item is or what do people need the most of. 

The dataset is from [Kaggle](https://www.kaggle.com/datasets/iamsouravbanerjee/customer-shopping-trends-dataset).


# Packages Used
```{r, message = FALSE}
#using this to tidy and filter the data with dplyr
library(tidyverse)
#using this to draw plots
library(ggplot2)
```


# Data at a glimpse
```{r, echo = FALSE}
shopping_data <- read_csv("shopping_trends.csv", show_col_types = FALSE)

shopping_data <- rename_with(shopping_data, ~ tolower(gsub(" ", "_", .x, fixed = TRUE)))

shopping_data <- rename(shopping_data, purchase_amount = 'purchase_amount_(usd)')


glimpse(shopping_data)

```
From just a glimpse we see a lot of columns of information. 


# Key Terms
- customer_id 
- age
- gender
- item_purchased
- category
- purchase_amount
- location
- size
- color
- season
- review_rating
- subscription_status
- payment_method
- shipping_type
- discount_applied
- promo_code_used
- previous_purchases
- preferred_payment_method
- frequency_of_purchases


# Questions
- How much are people spending? What are their ages?
- What gender is the main consumer?
- What sizes are in demand?
- What are they buying the most of, item and category?
- Is there a higher volume of sales in a specific state?
- Do consumers go for a specific color?
- What are buyers using as a way of payment?
- What are best/worst reviewed category?



# Items mainly bought:
```{r, echo = FALSE}
 
table_items <- shopping_data %>% 
  count(item_purchased) %>% 
  arrange(desc(n))

table_items


```

# Items least bought:
```{r, echo = FALSE}
 
table_items <- shopping_data %>% 
  count(item_purchased) %>% 
  arrange(n)

table_items


```

# What is the average of what people are spending for their age?
```{r, echo = FALSE}

table_age <- shopping_data %>% 
  count(age) %>% 
  arrange(desc(n))

table_age

shopping_data %>%
  group_by(age) %>%
  summarize(avg = mean(purchase_amount)) %>% arrange(desc(avg))

```


# What gender is the main consumer?
```{r, echo = FALSE}

shopping_data %>%
  group_by(gender) %>%
  summarize(avg = mean(purchase_amount))

```


# What sizes are in demand?
```{r, echo = FALSE}
table_size <- shopping_data %>% 
  count(size) %>% 
  arrange(desc(n))

table_size

```


# Is there a higher volume of sales in a specific state?
```{r, echo = FALSE}
 
table_location <- shopping_data %>% 
  count(location) %>% 
  arrange(desc(n))

table_location


```

# continued (least)
```{r, echo = FALSE}
 
table_location <- shopping_data %>% 
  count(location) %>% 
  arrange(n)

table_location


```


# Do consumers go for a specific color?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= color, fill = gender)) + facet_wrap(~ gender) + geom_bar() + labs(title = "Colors by gender", x = "Count", y = "Color") + theme(legend.position="none")
       
```

# What are they buying the most of?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= item_purchased, fill = gender)) + facet_wrap(~ gender) + geom_bar(position = "dodge") + labs(title = "Purchases by gender", x = "Count", y = "Item Purchased") + theme(legend.title=element_blank())
       
```


# Purchases by season
```{r, echo = FALSE}
ggplot(shopping_data, aes(purchase_amount, color = season)) +
  geom_freqpoly() + labs(title = "Purchases by season", x = "Purchase Amount", y = "Count") + theme(legend.title=element_blank())
       
```


# Purchases by category
```{r, echo = FALSE}

ggplot(shopping_data, aes(purchase_amount, color = category)) +
  geom_freqpoly() + labs(title = "Purchases by category", x = "Purchase Amount", y = "Count") + theme(legend.title=element_blank())
```


# What are buyers using as a way of payment?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(x = payment_method, fill = payment_method)) + geom_bar() + theme(legend.position="none") + labs(title = "Purchases by payment", x = "Payment Method", y = "Count")

```



# What categories have the best reviews?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= review_rating, x = category, fill = category)) + geom_boxplot() + labs(title = "Purchases by payment", x = "Review Rating", y = "Category") + theme(legend.title=element_blank())

```



# Summary

Understand your consumer.