---
title: "Shopping Trends"
author: "Marlen Gonzalez"
date: "2023-11-27"
output: slidy_presentation
---
# Introduction
I’ve always been the consumer but never looked at shopping from the point of view from the business side of it. I don’t usually go to a store and wonder what the main selling item is or what do people need the most of. I found this data set from Kaggle to open my eyes and learn a little bit more about what we as the consumer usually buy and who’s the main one buying things. 


# Packages Used
```{r}
#using this to tidy and filter the data with dplyr
library(tidyverse)
#using this to draw plots
library(ggplot2)
```


# Data at a glimpse
```{r, echo = FALSE}
shopping_data <- read_csv("shopping_trends.csv", show_col_types = FALSE)

shopping_data <- rename_with(shopping_data, ~ tolower(gsub(" ", "_", .x, fixed = TRUE)))

shopping_data <- rename(shopping_data, purchase_amount = 'purchase_amount_(usd)')


glimpse(shopping_data)

```
From just a glimpse we see a lot of columns of information. 

# Questions
- What age is the main consumer?
- What gender is the main consumer?
- What are they buying the most of, item and category?
- Is there a higher volume of sales in a specific state?
- What is the max amount people purchase?
- Do consumers go for a specific color?
- What are people mainly using to purchase items?
- What is the average of what people are spending for their age?

# Key Terms
- customer_id 
- age
- gender
- item_purchased
- category
- purchase_amount
- location
- size
- color
- season
- review_rating
- subscription_status
- payment_method
- shipping_type
- discount_applied
- promo_code_used
- previous_purchases
- preferred_payment_method
- frequency_of_purchases

# How much are the genders spending?
```{r, echo = FALSE}
 
table_items <- shopping_data %>% 
  count(item_purchased) %>% 
  arrange(desc(n))

table_items


```



# Ages
```{r, echo = FALSE}

table_age <- shopping_data %>% 
  count(age) %>% 
  arrange(desc(n))

table_age

shopping_data %>%
  group_by(age) %>%
  summarize(avg = mean(purchase_amount)) %>% arrange(desc(avg))

```
I was taken aback because I expected young adults to be the highest spenders.

# Gender
```{r, echo = FALSE}

shopping_data %>%
  group_by(gender) %>%
  summarize(avg = mean(purchase_amount))

```
I saw that coming but very close in average spending.

# Size 
```{r, echo = FALSE}
table_size <- shopping_data %>% 
  count(size) %>% 
  arrange(desc(n))

table_size

```

No surprise there, medium being the size mainly purchased!

# Location
```{r, echo = FALSE}
 
table_location <- shopping_data %>% 
  count(location) %>% 
  arrange(desc(n))

table_location


```

Montana is the main state buying in this data set.


# Color
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= color, fill = gender)) + geom_bar(position = "dodge")
       
```

# Item Purchased by gender
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= item_purchased, fill = gender)) + geom_bar(position = "dodge")
       
```
Most item bought is pants by men!

# Category by season
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= season, fill = category)) + geom_bar(position = "dodge")
       
```

The season people buy the most is spring and it's clothes!

# What a gender keeps coming back for more purchases
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= frequency_of_purchases, fill = gender )) + geom_bar(position = "dodge")
       
```
Every 3 months, men are the ones coming back to shop.


# Way of payment
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(x = payment_method)) + geom_bar()

```
Credit card was the mostly used form of payment

# What product has the best reviews
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= review_rating, x = category, fill = category))+ geom_boxplot()

```
This was just to see the difference in reviews


# Summary

From this data set it is surprisingly men who are shopping the most. If I was a retail store I would continue to market to that gender. 