---
title: "Shopping Trends"
author: "Marlen Gonzalez"
date: "2023-12-11"
output: html_document
---
# Introduction
I’ve always been a consumer but never looked at shopping from the point of view of the business side of it. I don’t usually go to a store and wonder what the popular selling item is or what people need the most. This dataset is AI generated but it was very useful to understand how a store would have to understand their consumers. 

The dataset is from [Kaggle](https://www.kaggle.com/datasets/iamsouravbanerjee/customer-shopping-trends-dataset).

The data set generated information for ages 18 and above and from all 50 states.

# Packages Used
```{r, message = FALSE}
#using this to tidy and filter the data with dplyr
library(tidyverse)
#using this to draw plots
library(ggplot2)
```


# Data at a glimpse
```{r, echo = FALSE}
shopping_data <- read_csv("shopping_trends.csv", show_col_types = FALSE)

shopping_data <- rename_with(shopping_data, ~ tolower(gsub(" ", "_", .x, fixed = TRUE)))

shopping_data <- rename(shopping_data, purchase_amount = 'purchase_amount_(usd)')


glimpse(shopping_data)

```
From just a glimpse we see a lot of columns of information. Fortunately the information is straightforward with very basic names and variables. The variables had spaces that would make it inconvenient to use in R. I renamed the variables to avoid the hassle.


# Key Terms
- customer_id 
- age
- gender
- item_purchased
- category
- purchase_amount
- location
- size
- color
- season
- review_rating
- subscription_status
- payment_method
- shipping_type
- discount_applied
- promo_code_used
- previous_purchases
- preferred_payment_method
- frequency_of_purchases

After looking through the data I started to see questions that I would have as a seller interested in making sales and trying to understand their consumer. 

- How much are people spending? What are their ages?
- What gender is the main consumer?
- What sizes are in demand?
- What are they buying the most of, item and category?
- Is there a higher volume of sales in a specific state?
- Do consumers go for a specific color?
- What are buyers using as a way of payment?
- What are best/worst reviewed category?



# Items mainly bought:
```{r, echo = FALSE}
 
table_items <- shopping_data %>% 
  count(item_purchased) %>% 
  arrange(desc(n))

table_items


```
Now here I started off slowly going through items mainly bought and I noticed that blouse jewelry pants were the main selling items.



# Items least bought:
```{r, echo = FALSE}
 
table_items <- shopping_data %>% 
  count(item_purchased) %>% 
  arrange(n)

table_items


```
While these items: backpacks gloves were the least sellers.  

# What is the average of what people are spending for their age?
```{r, echo = FALSE}

table_age <- shopping_data %>% 
  count(age) %>% 
  arrange(desc(n))

table_age

shopping_data %>%
  group_by(age) %>%
  summarize(avg = mean(purchase_amount)) %>% arrange(desc(avg))

```
The first table show the person’s age that spent the most at a store (69 spending ,88 dollars) while the other table shows on average people in their early 50s and late 40s were spending the most at more than 65 dollars.



# What gender is the main consumer?
```{r, echo = FALSE}

shopping_data %>%
  group_by(gender) %>%
  summarize(avg = mean(purchase_amount))

```

On the gender side of it females were on average the most to spend. I will note that the averages were very close! 

# What sizes are in demand?
```{r, echo = FALSE}
table_size <- shopping_data %>% 
  count(size) %>% 
  arrange(desc(n))

table_size

```
The most purchased size was Medium and least purchased size was Extra Large.

# Is there a higher volume of sales in a specific state?
```{r, echo = FALSE}
 
table_location <- shopping_data %>% 
  count(location) %>% 
  arrange(desc(n))

table_location


```

I was able to see what state was the main buyer and it was Montana. 


# continued (least)
```{r, echo = FALSE}
 
table_location <- shopping_data %>% 
  count(location) %>% 
  arrange(n)

table_location


```
The least states to buy was Kansas and Rhode Island.

# Do consumers go for a specific color?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= color, fill = gender)) + facet_wrap(~ gender) + geom_bar() + labs(title = "Colors by gender", x = "Count", y = "Color") + theme(legend.position="none")
       
```
I took into account that color is a main selling point in buying an Item but I also wanted see by gender so I graphed it classifying the colors and separating them by their genders like shown and I noticed that Silver was the main colors that men chose. While women chose yellow and I was also able to see the least color chosen by them which was brown, magenta and pink for men and turquoise and beige for women.


# What are they buying the most of?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= item_purchased, fill = gender)) + facet_wrap(~ gender) + geom_bar(position = "dodge") + labs(title = "Purchases by gender", x = "Count", y = "Item Purchased") + theme(legend.title=element_blank())
       
```
Continuing on gender I wanted to see what the most item purchased by each was, so I also did another graph showing that. Pants was by far the most purchased item by men. For women it was blouses. 

# Purchases by season
```{r, echo = FALSE, message= FALSE}
ggplot(shopping_data, aes(purchase_amount, color = season)) +
  geom_freqpoly() + labs(title = "Purchases by season", x = "Purchase Amount", y = "Count") + theme(legend.title=element_blank())
       
```
Next, I wanted to see the purchases by season so I did these graphs showing the purchase amount by the season I did notice that it was very similar so I tried doing another graph showing it by categories to see if there was any differences and I did see that there were more purchases on clothing compared to outerwear.

# Purchases by category
```{r, echo = FALSE, message= FALSE}

ggplot(shopping_data, aes(purchase_amount, color = category)) +
  geom_freqpoly() + labs(title = "Purchases by category", x = "Purchase Amount", y = "Count") + theme(legend.title=element_blank())
```


# What are buyers using as a way of payment?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(x = payment_method, fill = payment_method)) + geom_bar() + theme(legend.position="none") + labs(title = "Purchases by payment", x = "Payment Method", y = "Count")

```
From payment I was able to see that credit card was the main form of payment compared to bank transfer which in this day and age who does bank transfers. 


# What categories have the best reviews?
```{r, echo = FALSE}
ggplot(data = shopping_data, aes(y= review_rating, x = category, fill = category)) + geom_boxplot() + labs(title = "Purchases by payment", x = "Review Rating", y = "Category") + theme(legend.title=element_blank())

```
AND the last thing I wanted to see was just the reviews on items and what categories people were rating things on. I used the box plot to show the median and was able to see that clothing had the lowest rating. 


# Summary

Summary. Although it was AI generated I was able to gather information from the perspective of a retail seller. I was able to see the differences in what was the main selling point what colors were the most popular. I was able to see a lot of things that were would be interesting in the point of view of someone that was making money from understanding their consumer. This data set showed me if I was running a retail store or chain I would want my shops to have accessibility to credit card transactions or having a certain color like silver for men and yellow for women in stock, and having enough of medium sizes. In the end its just catering to your consumer. 